\begin{abstractpage}

Digital signatures applied in \href{"http://en.wikipedia.org/wiki/Cryptographic\_hash\_function\#Applications"}
{computer security}, are created and verified using cryptographic hash functions \cite{00042}.
The \href{"http://www.nist.gov/index.html"}{National Institute of Standards and Technology (NIST)} has published
family of cryptographic hash algorithms known as \href{http://en.wikipedia.org/wiki/Secure_Hash_Algorithm}
{Secure Hash Algorithm (SHA)} as U.S \href{http://en.wikipedia.org/wiki/Federal_Information_Processing_Standard}
{Federal Information Processing Standards (FIPS)}.

Due to advances in cryptanalysis of SHA-2, NIST announced a competition in November, 2007 to choose SHA-3. 
All the 64 submissions to NIST were open to public for cryptanalysis. 
In October, 2012 \href{"http://keccak.noekeon.org/"}{Keccak} was announced as the winner of SHA-3 competition. 
\href{"http://csrc.nist.gov/groups/ST/hash/sha-3/sha-3\_selection\_announcement.pdf"}{Keccak was chosen}
for its robustness and high security margin.

Of the 64 submissions for SHA-3 competition, 56 were selected for the first round, and 5 made it final round.
In this project I have compared the resistance to near collisions, on reduced versions of Keccak and two other
finalist \href{"https://131002.net/blake/"}{BLAKE}, and \href{"http://www.groestl.info/"}{Gr{\o}stl}. 
A $\epsilon / n $ bit near collision for hash function $h$ and two messages $M_{1}$ and $M_{2}$, where
$M_{1} \neq M_{2}$ is defined as
\begin{center}$HW( h( M_{1}, CV ) \oplus h( M_{2}, CV ) ) = n - \epsilon $\end{center}
where $HW$ is the Hamming weight, $h$ is the hash function, and $CV$ is the chaining value, and $n$ is the hash 
size in bits. I have also compared the near collision resistance amongst the different versions of Keccak by 
reducing its internal state size.

Hill climbing, simulated annealing, tabu search and random selection are applied to find a neighbourhood of chaining
value that is appended to two messages that are dissimilar by a bit, to find near collisions in the reduced versions 
of mentioned hashing algorithms.

Reduction in number of rounds predictably did reduce the security of hash function, but not by a large margin.
On other hand there was a inverse relation in increase in security margin and reduction in internal state size
of Keccak, for reduced rounds.
  
\end{abstractpage}
